---
title: "Action Potential - Group Analysis"
output: 
  html_document: default
  pdf_document:
    latex_engine: xelatex
    pandoc_args: --listings
date: "`r Sys.Date()`"
params:
  run: FALSE
  cores: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(doParallel)
library(foreach)
library(SobolSequence)
library(knitr)
source("ActionPotential.R")
source("vars.R")
```

```{r header, echo=FALSE}
#trace_files = c("Atratus_WT.csv", "Atratus_P.csv", "Atratus_EPN.csv")
group_file = "group_par"

with(as.list(par_0), {
  opt_par_group <<- c(#N_6=N_6,N_7=N_7,N_1=N_1,N_2=N_2,
                  M_6=M_6,M_7=M_7,M_1=M_1,M_2=M_2,
                  #H_4=H_4,H_5=H_5,H_6=H_6,H_3=H_3,
                  g_K=g_K, g_Na=g_Na, g_Leak=g_Leak)
})
```

```{r functions, echo=FALSE}
group_comp <- function(pars, vary_names, trace_d, time_d, id, cores) {
  num_fits = 10
  range = 0.25
  
  # Split the dataset
  vary = pars[names(pars) %in% vary_names]
  fixed = pars[! names(pars) %in% vary_names]

  # Create uniformly distributed set of fixed paramters
  varied_sobol = as.data.frame(sobolSequence.points(length(vary), count=num_fits)*(2*range) + (1-range))
  colnames(varied_sobol) = names(vary)
  for(r in 1:nrow(varied_sobol)) {
    varied_sobol[r,] = varied_sobol[r,] * vary
  }
  
  # Get varied experimental traces
  varied_traces = foreach(i = 1:num_fits, .combine = rbind) %do% {
    group_par = c(unlist(varied_sobol[i,]), fixed)
    AP_group = ActionPotential(group_par, trace_d, time_d)
    Vs_gen = AP_group$generate_trace(PLOT=TRUE)$Vs
    
    # Trim atrificially implied data generated from trace
    st = AP_group$stabil_time
    dt = AP_group$dt
    l = length(AP_group$trace_data)
    Vs_gen[(st/dt):(st/dt+l)]
  }
  
  # Optimize the original parameters againt the varied traces
  orig_param = c(fixed, vary)
  clust = makeCluster(cores)
  clusterExport(clust, c("varied_traces", "opt_par_group", "orig_param", "time_d"), envir=environment())
  registerDoParallel(clust)
  mixed_fits = foreach(i = 1:num_fits, .combine = rbind) %dopar% {
    source("ActionPotential.R")
    
    AP = ActionPotential(orig_param, unlist(varied_traces[i,]), time_d, "Group")
    out_par = AP$optimize(opt_par_group) #over everything
    
    data.frame(t(sapply(out_par$par,c)), values=out_par$value, convergence=out_par$convergence)
  }
  stopCluster(clust)
  
  saveRDS(mixed_fits, file = paste(output_folder, group_file, id, "fits.Rds", sep=""))
  saveRDS(varied_sobol, file = paste(output_folder, group_file, id, "sobol.Rds", sep=""))
  saveRDS(varied_traces, file = paste(output_folder, group_file, id, "traces.Rds", sep=""))
  
  return(mixed_fits)
}

show_group_comp <- function(id, time_d) {
  mixed_fits = readRDS(paste(output_folder, group_file, id, "fits.Rds", sep=""))
  varied_traces = readRDS(paste(output_folder, group_file, id, "traces.Rds", sep=""))
  
  num_fits = nrow(mixed_fits)
  for (i in 1:num_fits) {
    AP_group = ActionPotential(unlist(mixed_fits[i,]), varied_traces[i,], time_d)
    AP_group$display_action_potential()
  }
  
}

```

```{r main, echo=TRUE, results="asis"}
vary_names = c('g_Na', 'g_K', 'g_Leak')
if(params$run) {
  group_comp(par_0, vary_names, trace_data_ex, time_data_ex, "WT", params$cores)
}
show_group_comp("WT", time_data_ex)

# For each file -- to be used once working
if (FALSE) {
  # Running code
  for(file in trace_files) {
    data = read.csv(paste(data_folder, file, sep=""), skip=3)
    for(i in 1:(ncol(data)/2)) {
      name = names(data[2*i])
      trace_data = as.numeric(na.omit(data[,(2*i)]))
      time_data_loc = as.numeric(na.omit(data[,(2*(i-1)+1)]))
      fits = phi(fixed, vary, trace_data, time_data_loc, params$cores)
      saveRDS(fits, file = paste(output_folder_uni, file, name, group_file, ".Rds", sep=""))
    }
  }
  
  # Displaying code
  for(file in trace_files) {
  data = read.csv(paste(data_folder, file, sep=""), skip=3)
    for(i in 1:(ncol(data)/2)) {
      name = names(data[2*i])
      uni = readRDS(paste(output_folder_uni, file, name, unique_file, ".Rds", sep=""))
      print("Variable parameters:")
      print(vary_names)
      print(kable(uni, caption=paste("Fixed parameters uniformly spread using Sobel sequences for", name)))
    }
  }
}